Seismic Hazard Estimation of Northern Iran Using Smoothed Historical Seismicity

### 2016-01-31 Conclusion.tex
Many of Ricardo's comments were addressed.
Conclusion added.

\section{Conclusion}

The aim of this study is to conduct a new probabilistic seismic hazard assessment for northern Iran using updated seismic catalogs. Seismic hazard is evaluated over three regions (Azerbaijan, Tehran, Kopeh-Dagh) using the smoothed background seismicity. According to our study, in the Alborz  and Azerbaijan regions substantial seismic hazard is observed.  We studied probabilistic seismic hazard based on smoothed seismicity using historical and instrumental data. The results of this paper can be addressed in three main conclusions. First, probabilistic seismic hazard analysis using smoothed seismicity is sufficient for 10\% probability of exceedance. Second, since this study is based solely on historical and instrumental data, places with very close PGA values to those studies with considering faults as a primary source of seismicity, show normal and predictable seismic activity. Third, places with lower hazard values compared to other studies show anomalous seismic activity and suggest the idea of expecting more and bigger events in the future. As mentioned before, probabilistic seismic hazard analysis based on smoothed seismicity is very sensitive to input parameters (e.g., correlation distance, $b-value$, catalog completeness, and threshold magnitude for structural damage). Studying the sensitivity rate of these parameters could develop a more trustworthy result. Maps presented in this paper are for illustrative purposes only, and are not intended to be used in any application.

>>>>

\section{Conclusion}
The result of a new probabilistic seismic hazard assessment for northern Iran using updated seismic catalogs is provided. Seismic hazard is evaluated over entire three regions (Azerbaijan, Alboz, Kopeh-Dagh) as well as part of Central and Zagros tectonic seismic regions using the smoothed background seismicity. In order to determine the seismic hazard in North Iran, different studies have been done. These studies are based on defining faults as a seismic source model. Using faults as a seismic source model could result in accurate values in case the faults are precisely determined. There are considerable uncertainties on fault characteristic, specially on fault length. Instead of using faults as a seismic source, we consider each earthquake event as a point source in the model. We reduce the uncertainty in earthquake location by smoothing the number of earthquake in adjacent cells. We defined models based on considering the whole North Iran as a uniform seismic tectonic regions and 5 different seismic tectonic regions. The differences in catalog completeness, b value, and maximum magnitude of earthquake in these two models resulted in different results. This indicate the idea that dividing regions into subregions is necessary in order to be able to get accurate results. The result are also sensitive to the minimum magnitude. The structures built according to engineering code are supposed to withstand the earthquake with magnitude 5 at the reasonable distance. However, closer $M_w > 4.5$ for those structures and also older masonry structures could cause structural damage. A combination of two models $(M_w > 4.5, M_w>5)$ in each city based on number of engineering and old masonry structures will provide more accurate results. The results are provided for mean peak horizontal acceleration in bedrock. The ground acceleration in soil deposits and inside basin will be bigger than the provided values. This model reduce the epistemic uncertainty regarding the recognizing faults and earthquake location. Combination of this model with a study based on using faults will increase the values for existing non-active seismic fault regions.

### 2016-01-31 abstract.tex
Conclusion added.

In this study we conduct a new probabilistic seismic hazard assessment for northern Iran, using recently confirmed ground motion prediction equations and updated seismic catalogs. We evaluate seismic hazard over whole three tectonic seismic zones (Alborz Mountain Range, Azerbaijan and Kopeh-Dagh) and part of two tectonic seismic zones (Zagros and Central-East Iran) using smoothed historical seismicity. In this model, events are not assigned to specific faults and are assumed to be potential seismogenic sources, spatially gridded to cells. We decluster historical and instrumental seismicity to compute the rate of earthquakes in each grid, and smooth these rates to account for uncertainty in the spatial distribution of future earthquakes. We assume each grid cell has spacing of $0.1^{\circ}$ in latitude and $0.1^{\circ}$ in longitude and count the number of earthquakes in each of them. We use Gaussian filter to smooth the fraction of earthquakes in incremental intervals of magnitude. For this article, in order to consider the seismicity rate, we consider all historical and instrumental earthquakes with magnitude above 3. To calculate the seismic hazard we defined two models based on structural damage threshold magnitude, including Mw 4.5 and Mw 5. Seismicity parameters are calculated for each seismic region, and smoothed seismicity is calculated separately for each. We also consider the combination of regions as a uniform region.  We add the probability of exceedance of all regions and present the smoothed seismicity hazard in northern Iran. The horizontal peak ground acceleration is selected as the ground motion parameter. The results are assessed in terms of the expected peak ground acceleration (PGA) with a 10\% and 2\% probability of exceedance in 50 years for rock site conditions. The seismic hazard map for northern Iran is presented. \textcolor{red}{The maximum seismic hazard values are 0.44 $g$ and  0.74 $g$ for 10\% and 2\% probability in exceedance in 50 years, which are calculated in both Azerbaijan and Alborz tectonic seismic region. The result of 10\% probability of exceedance in 50 years are in close agreement with current Iranian code of practice for seismic design of building.}

>>>>

In this study we conduct a new probabilistic seismic hazard assessment for northern Iran, using recently confirmed ground motion prediction equations and updated seismic catalogs. We evaluate seismic hazard over whole three tectonic seismic zones (Alborz Mountain Range, Azerbaijan and Kopeh-Dagh) and part of two tectonic seismic zones (Zagros and Central-East Iran) using smoothed seismicity. In this model, events are not assigned to specific faults and are assumed to be potential seismogenic sources, spatially gridded to cells. We decluster historical and instrumental seismicity to compute the rate of earthquakes in each grid, and smooth these rates to account for uncertainty in the spatial distribution of future earthquakes. We assume each grid cell has spacing of $0.1^{\circ}$ in latitude and $0.1^{\circ}$ in longitude and count the number of earthquakes in each of them. We use Gaussian filter to smooth the fraction of earthquakes in incremental intervals of magnitude. For this article, in order to consider the seismicity rate, we consider all historical and instrumental earthquakes with magnitude above 3. To calculate the seismic hazard we defined two models based on structural damage threshold magnitude, including Mw 4.5 and Mw 5. Seismicity parameters are calculated for each seismic region, and smoothed seismicity is calculated separately for each. We also consider the combination of regions as a uniform region.  We add the probability of exceedance of all regions and present the smoothed seismicity hazard in northern Iran. The horizontal peak ground acceleration is selected as the ground motion parameter. The results are assessed in terms of the expected peak ground acceleration (PGA) with a 10\% and 2\% probability of exceedance in 50 years for rock site conditions. The seismic hazard map for northern Iran is presented.

### 2016-01-31 analysis_parameters.tex
working on discussion and conclusion part. Adding annual rate of earthquake figure.
Conclusion added.

The attenuation relationship is:

@@

For each of the three seismic regions, Gutenberg-Richter parameters and Max magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in each grid cells.\\
\noindent
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the north Iran uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of this study \citep{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  
\noindent

@@

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.6]{figures/pdf/completness_scatter.pdf} 
\caption{Magnitude-time distribution of earthquakes in the study regions. a) Alborz, b) Azerbaijan, c)Central, d)Kopeh Dagh, e)Zagros}
\label{fig:completness_scatter}
\end{figure*}

 According to   \citet{Frankel1995},  we made plots of the cumulative number of events against time for different regions. We pick the Mw 0.5 increments in magnitude to be able to compare the results with \citet{Zare2014}. In order to be able to compare the results with \citet{Zare2014} we merge the data of Azerbaijan and Alborz seismic regions.  Fig.~\ref{fig:completness_compare_zare_2014_Az_Al} shows the completeness of catalog for earthquakes with different magnitude range in Azerbaijan-Alborz region. According to this figure, midrange magnitudes ($ 4 < Mw < 6 $) fairly obey the network developments in 1900 and 1961. The completeness thresholds for each magnitude range which are reported by \citet{Zare2014} are shown by dashed green line. In this study we follow the \citet{Frankel1995} approach to determine the completeness of each region. Even though the approach used by \citet{Frankel1995} will lead to the conservative results, we make sure that we don't use a period of time without knowing the complete number of events. Determining the completeness of the catalog is very sensitive to the data. Converting earthquake magnitudes from different scales to moment magnitude obviously has some error. Having broader range of magnitude will help to minimize these sort of error. In this study we consider the magnitude intervals for completeness study as $Mw = 1$.

@@

\noindent
Fig.~\ref{fig:comp_test_all_mag} illustrates the completeness of data for each magnitude threshold in five different regions. A uniform rise of cumulative number of earthquake in each magnitude range, defines the threshold for catalog completeness. We define a completeness of each magnitude range at a time which the cumulative number of earthquake increase linearly with time. We assume the catalog for earthquakes with magnitude greater than  7 ($M_w > 7$) is complete from the first historical earthquake report.  Determining the completeness of the catalog for $ 6 < Mw \leq 7 $ needs precise observation of the catalog. In Azerbaijan tectonic seismic region after 1045  $Mw = 6.2$ earthquake the catalog seems complete up to 1318. However, between 1318 and 1581 (263 years) only one earthquake with magnitude $Mw > 6$ is reported (1440,  $Mw = 6.2$ ). Considering the activity of the Azerbaijan region, in this study we assume that the catalog is complete for earthquake with magnitude $ 6 < Mw \leq 7 $ after 1581. Similar situation happened at the period of 1715 - 1833. It worth to mention that , the time period is shorter than before (118 years), meanwhile in this period of time two earthquakes with magnitude $Mw>7$ occurred in the Azerbaijan region (1721 $Mw=7.7$, 1780 $Mw=7.6$). The catalog may or may not be completed after 1045, however in this study we consider 1581 as a completeness threshold for magnitude  $ 6 < Mw \leq 7 $. Considering completeness of the catalog for Azerbaijan region from 1581 will result in conservative PGA for the region in compare with the model with threshold of 1045. We picked the completeness threshold for other regions and other magnitude ranges according as a year that the cumulative number of earthquakes increment uniformly. Table ~\ref{tab:completeness} shows the summary of values used in this study.

>>>>

\noindent
The attenuation relationship is:

@@

For each of the three seismic regions, Gutenberg-Richter parameters and Max magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in each grid cells.\\
\noindent
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog.  Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, 4.5, 4.5, and 4.4 for Azerbaijan, Alborz,  Kopeh-Dagh, Central Iran, and Zagros tectonic seismic regions, respectively. For the north Iran uniform model we used 4.5 as a completeness magnitude. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of this study \citep{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table \ref{tab:b_value}.  Annual occurrence rate of the earthquake for each region is shown in Fig.~\ref{fig:annual_m}.

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.4]{figures/pdf/annual_rate_m.pdf} 
\caption{Annual occurrence rate of the earthquake. a) Azerbaijan, b) Alborz, c) KopehDagh, d)Central, e)Zagros f) North Iran}
\label{fig:annual_m}
\end{figure*}

@@

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.6]{figures/pdf/completness_scatter.pdf} 
\caption{Magnitude-time distribution of earthquakes in the study regions. a) Alborz, b) Azerbaijan, c)Central, d)Kopeh-Dagh, e)Zagros}
\label{fig:completness_scatter}
\end{figure*}

 According to   \citet{Frankel1995},  we made plots of the cumulative number of events against time for different regions' catalog. We pick the Mw 0.5 increments in magnitude to be able to compare the results with \citet{Zare2014}. In order to be able to compare the results with \citet{Zare2014} we merge the data of Azerbaijan and Alborz seismic regions.  Fig.~\ref{fig:completness_compare_zare_2014_Az_Al} shows the completeness of catalog for earthquakes with different magnitude range in Azerbaijan-Alborz region. According to this figure, midrange magnitudes ($ 4 < Mw < 6 $) fairly obey the network developments in 1900 and 1961. The completeness thresholds for each magnitude range which are reported by \citet{Zare2014} are shown by dashed green line. In this study we follow the \citet{Frankel1995} approach to determine the completeness of each region. Even though the approach used by \citet{Frankel1995} will lead to the conservative results, we make sure that we don't use a period of time without knowing the complete number of events. Determining the completeness of the catalog is very sensitive to the data. Converting earthquake magnitudes from different scales to moment magnitude obviously has some error. Having broader range of magnitude will help to minimize these sort of error. In this study we consider the magnitude intervals for completeness study as $Mw = 1$.

@@

\noindent
Fig.~\ref{fig:comp_test_all_mag} illustrates the completeness of data for each magnitude threshold in five different regions. A uniform rise of cumulative number of earthquake in each magnitude range, defines the threshold for catalog completeness. We define the completeness of each magnitude range at a time which the cumulative number of earthquake increase linearly with time. We assume the catalog for earthquakes with magnitude greater than  7 ($M_w > 7$) is complete from the first historical earthquake report.  Determining the completeness of the catalog for $ 6 < Mw \leq 7 $ needs precise observation of the catalog. In Azerbaijan tectonic seismic region after 1045  $Mw = 6.2$ earthquake the catalog seems complete up to 1318. However, between 1318 and 1581 (263 years) only one earthquake with magnitude $Mw > 6$ is reported (1440,  $Mw = 6.2$ ). Considering the activity of the Azerbaijan region, in this study we assume that the catalog is complete for earthquake with magnitude $ 6 < Mw \leq 7 $ after 1581. Similar situation happened at the period of 1715 - 1833. It worth to mention that in this case, the time period is shorter than before (118 years), meanwhile in this period of time two earthquakes with magnitude $Mw>7$ occurred in the Azerbaijan region (1721 $Mw=7.7$, 1780 $Mw=7.6$). The catalog may or may not be completed after 1045, however in this study we consider 1581 as a completeness threshold for magnitude  $ 6 < Mw \leq 7 $. Considering completeness of the catalog for Azerbaijan region from 1581 will result in conservative PGA for the region in compare with the model with threshold of 1045. We picked the completeness threshold for other regions and other magnitude ranges according as a year that the cumulative number of earthquakes increment uniformly. Table ~\ref{tab:completeness} shows the summary of values used in this study.

### 2016-01-31 introduction.tex
Conclusion added.

\section{Introduction}
Iran is situated over Himalayan-Alpide seismic belt, which has frequently experienced strong shaking induced by earthquakes. The occurrence of these earthquakes has imposed notable destruction to the buildings and lifelines, and unfortunately, has caused extensive loss of human life. These earthquakes are categorized in various tectonic seismic zones. Different tectonic seismic regions have been defined for Iran in different studies.  Some of these studies defined more detailed division \citep{Nowroozi1976, Tavakoli1999}, and some of them defined one simplified province \citep{Stocklin1968, Takin1972, Berberian1976}. \citet{Mirzaei1998} divided Iran into five tectonic regions, including Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran. Regarding the other tectonic seismic regions, \citet{Karimiparidari2013} divided the Azerbaijan-Alborz tectonic seismic region into two regions, which are Azerbaijan and Alborz Mountain Range (hereinafter, Alborz).\\
 Azerbaijan, Alborz and Kopeh-Dagh tectonic seismic regions encompass most of northern Iran. Northern Iran has many highly populated cities (e.g., Tehran, Tabriz, and Mashhad), in which many destructive earthquakes have been reported. \\
\noindent
In the last three decades, different seismic hazard analysis studies \citep[e.g.,][]{Tavakoli1999,Ghodrati2003} substantially improved the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}.
\citet{Ghodrati2013} conducted a seismic risk assessment for the city of Tehran using the Hazus method \citep{FEMA2003}. They represent that these efforts successfully decrease the total mean damage ratio in Tehran from 0.302739 in 1996 to 0.272859 in 2006.  Even though the results are satisfactory, the continuous improvement of procedures for defining the seismic hazard at regional (national) and local levels is essential for the optimum design of earthquake-resistant structures. In this study we use another model to overcome the aleatory and epistemic uncertainties regarding earthquake location and source mechanism in northern Iran.\\

In this study we used smoothed seismicity approach \citep{Frankel1995} in order to do seismic hazard analysis. We divided northern Iran in three seismotectonic regions including Azerbaijan, Alborz, and Kopek Dagh. We also consider seismicity of part of Central-East Iran and Zagros regions. We computed seismic parameters and catalog completeness in these regions. Zagros and Central Iran seismotectonic regions have considerable effect on the Northern part.   We consider the $Mw \  5$ earthquake as a threshold magnitude  for structural damage. In order to consider damage to the historical masonry building (which includes most part of big cities and rural areas) we also considered $Mw \  4.5$ as a threshold for structural damage to these kind of building.

>>>>

\section{Introduction}
Iran is situated over Himalayan-Alpide seismic belt, which has frequently experienced strong shaking induced by earthquakes. The occurrence of these earthquakes has imposed notable destruction to the buildings and lifelines, and unfortunately, has caused extensive loss of human life. These earthquakes are categorized in various tectonic seismic zones. Different tectonic seismic regions have been defined for Iran in different studies.  Some of these studies defined more detailed division \citep{Nowroozi1976, Tavakoli1999}, and some of them defined one simplified province \citep{Stocklin1968, Takin1972, Berberian1976}. \citet{Mirzaei1998} divided Iran into five tectonic regions, including Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran. Regarding the other tectonic seismic regions, \citet{Karimiparidari2013} divided the Azerbaijan-Alborz tectonic seismic region into two regions, which are Azerbaijan and Alborz Mountain Range (hereinafter, Alborz).\\
 Azerbaijan, Alborz and Kopeh-Dagh tectonic seismic regions encompass most of northern Iran. Northern Iran has many highly populated cities (e.g., Tehran, Tabriz, and Mashhad), in which many destructive earthquakes have been reported. \\
\noindent
In the last three decades, different seismic hazard analysis studies \citep[e.g.,][]{Tavakoli1999,Ghodrati2003} substantially improved the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}.
\citet{Ghodrati2013} conducted a seismic risk assessment for the city of Tehran using the HAZUS method \citep{FEMA2003}. They represent that these efforts successfully decrease the total mean damage ratio in Tehran from 0.302739 in 1996 to 0.272859 in 2006.  Even though the results are satisfactory, the continuous improvement of procedures for defining the seismic hazard at regional (national) and local levels is essential for the optimum design of earthquake-resistant structures. In this study we use another model to overcome the aleatory and epistemic uncertainties regarding earthquake location and source mechanism in northern Iran.\\

In this study we used smoothed seismicity approach \citep{Frankel1995} in order to do seismic hazard analysis. We divided northern Iran in three seismotectonic regions including Azerbaijan, Alborz, and Kopeh Dagh. We also consider seismicity of part of Central-East Iran and Zagros regions. We computed seismic parameters and catalog completeness in these regions. Zagros and Central Iran seismotectonic regions have considerable effect on the Northern part.   We consider the $Mw \  5$ earthquake as a threshold magnitude  for structural damage. In order to consider damage to the historical masonry building (which includes most part of big cities and rural areas) we also considered $Mw \  4.5$ as a threshold for structural damage to these kind of building.
